**The status of the microbial census: an update**

### 2015-11-18 Schloss_Census2_mBio_2015.Rmd
Fix some typos

bact <- read.table(file='data/process/bacteria.v123.metadata', header=T, row.names=1, stringsAsFactors=FALSE)

@@

* Why aren't we doing a great job?
  - Rate of change in full-length sequences has plateaued
  - New sequences are submitted by a small number of investigators for a limited number of environments

@@

	require(knitcitations, quiet=TRUE)
	citep(c('10.1038/ismej.2012.79', '10.1371/journal.pone.0026284', '10.1101/gr.131029.111', '10.1073/pnas.1004204107', '10.1126/science.1171700'))

@@

***The dynamics of the bacterial census.***
One explanation for the large number of OTUs that have only been observed once is that with the the broad adoption of highly parallelized sequencing platforms that generate short sequence reads, the rate of full-length sequence generation has declined. In fact, since 2009 the number of new bacterial sequences generated has plateaued at an average of `r mean(time_data[as.character(2009:2014),'b_nseqs'])` sequences per year (Figure 2A). Although this is still an impressive number of sequences, since 2008 the number of new bacterial OTUs has plateaued at an average of `r mean(time_data[as.character(2006:2014),'b_sobs'])` new OTUs per year (Figure 2B). Given the expense of generating full-length sequences using the Sanger sequencing technology and the transition to other platforms, we expected that the large number of sequences were being deposited by a handful of large projects. Indeed, when we counted the number of submissions responsible for depositing 50% of the sequences, we found that with the exception of 2006 and 2013, eight or fewer studies were responsible for depositing the majority of the full-length sequences each year since 2005 (Figure 2C). Between 2009 and 2012, `r sum(time_data[as.character(2009:2012), 'b_nseqs'])` total sequences were submitted and `r sum(time_data[as.character(2009:2012), 'b_n50'])`) submissions from `r length(unique(dominant_titles))` studies were responsible for depositing `r dominant_studies` (`r 100*dominant_studies/sum(time_data[as.character(2009:2012), 'b_nseqs'])`% of all sequences). These studies generated sequences from the human gastrointestinal tract [@Li_2012], human skin [@Kong_2012; @Grice_2009], murine skin [@Grice_2010], and hypersaline microbial mats [@Kirk_Harris_2012]. In contrast to recent years, between 1995 and 2006, an average of `r mean(time_data[as.character(1995:2006), 'b_n50'])` studies were responsible for submitting more than half of the sequences each year. Although these deep surveys represent significant contributions to our knowledge of bacterial biogeography, their small number and lack of environmental diversity is indicative of the broader problems in advancing the bacterial census.

>>>>

bact <- read.table(file='data/process/bacteria.v123.metadata', header=T, row.names=1, stringsAsFactors=FALSE)
bact_year_deposited <- gsub("^(\\d\\d\\d\\d)-.*", "\\1", bact$date)

@@

@@

	#require(knitcitations, quiet=TRUE)
	#citep(c('10.1038/ismej.2012.79', '10.1371/journal.pone.0026284', '10.1101/gr.131029.111', '10.1073/pnas.1004204107', '10.1126/science.1171700'))

@@

***The dynamics of the bacterial census.***
One explanation for the large number of OTUs that have only been observed once is that with the the broad adoption of highly parallelized sequencing platforms that generate short sequence reads, the rate of full-length sequence generation has declined. In fact, since 2009 the number of new bacterial sequences generated has plateaued at an average of `r floor(mean(time_data[as.character(2009:2014),'b_nseqs']))` sequences per year (Figure 2A). Although this is still an impressive number of sequences, since 2008 the number of new bacterial OTUs has plateaued at an average of `r mean(time_data[as.character(2006:2014),'b_sobs'])` new OTUs per year (Figure 2B). Given the expense of generating full-length sequences using the Sanger sequencing technology and the transition to other platforms, we expected that the large number of sequences were being deposited by a handful of large projects. Indeed, when we counted the number of submissions responsible for depositing 50% of the sequences, we found that with the exception of 2006 and 2013, eight or fewer studies were responsible for depositing the majority of the full-length sequences each year since 2005 (Figure 2C). Between 2009 and 2012, `r sum(time_data[as.character(2009:2012), 'b_nseqs'])` total sequences were submitted and `r sum(time_data[as.character(2009:2012), 'b_n50'])` submissions from `r length(unique(dominant_titles))` studies were responsible for depositing `r dominant_studies` (`r 100*dominant_studies/sum(time_data[as.character(2009:2012), 'b_nseqs'])`% of all sequences). These studies generated sequences from the human gastrointestinal tract [@Li_2012], human skin [@Kong_2012; @Grice_2009], murine skin [@Grice_2010], and hypersaline microbial mats [@Kirk_Harris_2012]. In contrast to recent years, between 1995 and 2006, an average of `r mean(time_data[as.character(1995:2006), 'b_n50'])` studies were responsible for submitting more than half of the sequences each year. Although these deep surveys represent significant contributions to our knowledge of bacterial biogeography, their small number and lack of environmental diversity is indicative of the broader problems in advancing the bacterial census.

==> The archaea census section is pretty rough still...
